Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-01-28 paper/compilation.tex
fix label reference issue
more edits

\sysname also supports constraints purely on the control-plane behavior of BGP. For example, prefix aggregation, which should not affect routing behavior, is an important optimization to reduce routing table size and churn in practice. 
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but we also support other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

@@

The semantics of RIR is in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \Prefer \dots \Prefer r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstract locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$ and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the length of the path. Lower ranks indicate \emph{more} preferred paths. Traffic may be sent along any of the most preferred paths for each pair of starting and ending locations that appear in some valid specified path.

@@

The main differences between the FE and RIR are: (1) FE allows the programmer to specify constraints using a series of (modular) definitions, and combine them later, (2) FE provides high-level names that abstract sets of routes, and (3) FE allows the preference operator to be used locally.

A key constraint when translating FE to RIR is to ensure that all specified routes are well-formed. In particular, each regular path constraint $r$ must satisfy $r \subseteq \Out^* \cdot \In^+ \cdot \Out^*$. This ensures that one can only talk about controlling traffic that goes through the user's network at some point,
and that such traffic does not loop back multiple times to the user network.

@@

The next step rewrites the high-level constraints such as \textit{enter} according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing equivalence:
%
$$x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n)$$
%
where $\odot$ stands for an arbitrary regular binary operator. Preferences nested under a unary operator, \textit{star} or \textit{negation}, are flagged by the compiler as invalid policies.

@@

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{in}{}
    \transition{2}{2}{A,C,D,E}{loop above}
    \transition{2}{3}{B}{}
    \transition{3}{3}{B}{loop above}
    \transition{3}{2}{A,C,D,E}{bend left}
    \transition{3}{4}{W}{}
  \end{tikzpicture}

  \end{minipage}
  %
  ~~
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Product Graph IR}{}
  \vspace*{-1\baselineskip}
  %
  \includegraphics[width=.65\columnwidth]{figures/productgraph}
  \end{minipage}

@@

Before checking if policies captured by the PGIR are safe, we minimize the PGIR to improve our safety analysis. 
The minimization is based on the observation that, although every path in the PGIR is a valid path in the topology, we do not want to consider paths that form loops. In particular, BGP's loop prevention mechanism forces an AS to reject any route for which it is already in the AS path.
%
For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
It often simplifies the failure-safety analysis to eliminate cases that are not possible due to loops. 
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.
We use graph dominators~\ref{bib:todo} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading to $y$ from the start node. Many efficient algorithms exist for finding graph dominators in practice~\ref{bib:todo}.
%We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node. 
For example, node $(1,1) W$ in Figure~\ref{fig:example-compilation} is never on a simple path to the end node since it most go through another $W$ in either case.
Colored nodes and dashed edges are removed after minimization since they are irrelvant to the BGP decision process.
%
%Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

@@

\para{Avoiding Simple Paths}

@@

\begin{figure}[t!]
%% \begin{lstlisting}[frame=single, mathescape=true]
%% $\Router{A}$
%%   Match $\PEER=C, \COMM=(3,2)$
%%     Export $\COMM \leftarrow (4,2), \MED \leftarrow 80, \PEER \leftarrow W$
%% $\Router{B}$
%%   Match $\PEER = C, \COMM = (-,2)$
%%     Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
%%            $\MED \leftarrow 81, \PEER \leftarrow W$
%%   Match $\PEER = C, \COMM = (3,2)$
%%     Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
%%            $\MED \leftarrow 81, \PEER \leftarrow W$
%% $\Router{C}$
%%   Match$[LP=99]$ $\PEER = E, \COMM = (-,2) $
%%     Export $\COMM \leftarrow (-,2), \PEER \leftarrow B$
%%   Match $\PEER = D, \COMM = (2,2)$
%%     Export $\COMM \leftarrow (3,2), \PEER \leftarrow A,B$
%% $\Router{D}$
%%   Match $\REGEX{X + Y}$
%%     Export $\COMM \leftarrow (2,2), \PEER \leftarrow C$
%% $\Router{E}$
%%   Match $\REGEX{Z}$
%%     Export $\COMM \leftarrow (-,2), \PEER \leftarrow C$
%% \end{lstlisting}
\begin{code}
\Router{A}
  Match \PEER=C, \COMM=(3,2)
    Export \COMM \Arrow (4,2), \MED \Arrow 80, \PEER \Arrow W
\Router{B}
  Match \PEER = C, \COMM = (-,2)
    Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
           \MED \Arrow 81, \PEER \Arrow W
  Match \PEER = C, \COMM = (3,2)
    Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
           \MED \Arrow 81, \PEER \Arrow W
\Router{C}
  Match[LP=99] \PEER = E, \COMM = (-,2) 
    Export \COMM \Arrow (-,2), \PEER \Arrow B
  Match \PEER = D, \COMM = (2,2)
    Export \COMM \Arrow (3,2), \PEER \Arrow A,B
\Router{D}
  Match \REGEX{X + Y}
    Export \COMM \Arrow (2,2), \PEER \Arrow C
\Router{E}
  Match \REGEX{Z}
    Export \COMM \Arrow (-,2), \PEER \Arrow C
\end{code}\vspace{-.5cm}
\label{fig:abgp-config}
\caption{Abstract BGP router configurations.}
\end{figure}

>>>>

\sysname also supports constraints purely on the control-plane behavior of BGP. For example, prefix aggregation is an important optimization to reduce routing table size and churn in practice. 
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but we also support other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

@@

The semantics of RIR is in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \Prefer \dots \Prefer r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstract locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$, that is the path is in the language of the regular expression, and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the length of the path. Lower ranks indicate \emph{more} preferred paths. Traffic may be sent along any of the most preferred paths for each pair of starting and ending locations that appear in some valid specified path.

@@

The main differences between the FE and RIR are: $i)$ FE allows the programmer to specify constraints using a series of (modular) definitions, and combine them later, $ii)$ FE provides high-level names that abstract sets of routes and groups of prefixes/neighbors, and $iii)$ FE allows the preference operator to be used more flexibly.

A key constraint when translating FE to RIR is to ensure that all specified routes are well-formed. In particular, each regular path constraint $r$ must satisfy $r \subseteq \Out^* \cdot \In^+ \cdot \Out^*$. This ensures that the user only control traffic that goes through the user's network at some point, and that such traffic does not loop back multiple times to the user network.

@@

The next step rewrites the high-level constraints such as \textit{enter} according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing distributivity equivalences:
%
\[ 
\begin{array}{c}
  x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n) \\
  (y_1 \Prefer \dots \Prefer y_n) \odot x = (y_1 \odot x) \Prefer \dots \Prefer (y_n \odot x)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator. Preferences nested under a unary operator, \textit{star} or \textit{negation}, are flagged by the compiler as invalid policies.

@@

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{in}{}
    \transition{2}{2}{A,C,D,E}{loop above}
    \transition{2}{3}{B}{}
    \transition{3}{3}{B}{loop above}
    \transition{3}{2}{A,C,D,E}{bend left}
    \transition{3}{4}{W}{}
  \end{tikzpicture}
  \end{minipage}
  %
  ~~
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Product Graph IR}{}
  \vspace*{-1\baselineskip}
  %
  \includegraphics[width=.65\columnwidth]{figures/productgraph}
  \end{minipage}

  \hrulefill
  \vspace*{.4em}

@@

Before checking if policies captured by the PGIR are safe, we minimize the PGIR to improve our safety analysis. 
The minimization is based on the observation that, although every path in the PGIR is a valid path in the topology, we do not want to consider paths that form loops. In particular, BGP's loop prevention mechanism forces an AS to reject any route for which it is already in the AS path.
%
For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
It often simplifies the failure-safety analysis to eliminate cases that are not possible due to loops. 
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.
We use graph dominators~\ref{bib:todo} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading to $y$ from the start node. Many efficient algorithms exist for finding graph dominators in practice~\ref{bib:todo}.
%We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node. 
For example, node $(1,1) W$ in Figure~\ref{fig:example-compilation} is never on a simple path to the end node since it must go through another $W$ in either case.
Colored nodes and dashed edges are removed after minimization since they are irrelvant to the BGP decision process.
%
%Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

@@

\para{Avoiding Loops}

@@

\begin{figure}[t!]
%% \begin{lstlisting}[frame=single, mathescape=true]
%% $\Router{A}$
%%   Match $\PEER=C, \COMM=(3,2)$
%%     Export $\COMM \leftarrow (4,2), \MED \leftarrow 80, \PEER \leftarrow W$
%% $\Router{B}$
%%   Match $\PEER = C, \COMM = (-,2)$
%%     Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
%%            $\MED \leftarrow 81, \PEER \leftarrow W$
%%   Match $\PEER = C, \COMM = (3,2)$
%%     Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
%%            $\MED \leftarrow 81, \PEER \leftarrow W$
%% $\Router{C}$
%%   Match$[LP=99]$ $\PEER = E, \COMM = (-,2) $
%%     Export $\COMM \leftarrow (-,2), \PEER \leftarrow B$
%%   Match $\PEER = D, \COMM = (2,2)$
%%     Export $\COMM \leftarrow (3,2), \PEER \leftarrow A,B$
%% $\Router{D}$
%%   Match $\REGEX{X + Y}$
%%     Export $\COMM \leftarrow (2,2), \PEER \leftarrow C$
%% $\Router{E}$
%%   Match $\REGEX{Z}$
%%     Export $\COMM \leftarrow (-,2), \PEER \leftarrow C$
%% \end{lstlisting}
\begin{code}
\Router{A}
  Match \PEER=C, \COMM=(3,2)
    Export \COMM \Arrow (4,2), \MED \Arrow 80, \PEER \Arrow W
\Router{B}
  Match \PEER = C, \COMM = (-,2)
    Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
           \MED \Arrow 81, \PEER \Arrow W
  Match \PEER = C, \COMM = (3,2)
    Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
           \MED \Arrow 81, \PEER \Arrow W
\Router{C}
  Match[LP=99] \PEER = E, \COMM = (-,2) 
    Export \COMM \Arrow (-,2), \PEER \Arrow B
  Match \PEER = D, \COMM = (2,2)
    Export \COMM \Arrow (3,2), \PEER \Arrow A,B
\Router{D}
  Match \REGEX{X + Y}
    Export \COMM \Arrow (2,2), \PEER \Arrow C
\Router{E}
  Match \REGEX{Z}
    Export \COMM \Arrow (-,2), \PEER \Arrow C
\end{code}\vspace{-.5cm}
\caption{Abstract BGP router configurations. \label{fig:abgp-config}}
\end{figure}
